---
title: "Getting Started with plsRglm"
author:
  - "Frederic Bertrand"
  - "Myriam Maumy-Bertrand"
date: "March 2026"
output:
  rmarkdown::html_vignette:
    toc: true
    toc_depth: 2
vignette: >
  %\VignetteIndexEntry{Getting Started with plsRglm}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  message = FALSE,
  warning = FALSE,
  fig.width = 7,
  fig.height = 5,
  dpi = 150
)

set.seed(123)
library(plsRglm)
```

`plsRglm` provides partial least squares regression for linear and generalized linear models, repeated k-fold cross-validation, bootstrap utilities, and support for incomplete predictor matrices. This vignette is the practical starting point for the current package API. The companion vignette `vignette("plsRglm", package = "plsRglm")` keeps the longer historical case studies and algorithmic notes.

# Core Fitting Workflows

`plsR()` is the dedicated interface for ordinary PLS regression. `plsRglm()` extends the same ideas to generalized linear and ordinal models, and can also fit `modele = "pls"` through the shared interface.

## Linear PLS with matrix and formula interfaces

```{r linear-pls}
data(Cornell)
XCornell <- Cornell[, 1:7]
yCornell <- Cornell$Y

pls_fit_matrix <- plsR(yCornell, XCornell, nt = 3, verbose = FALSE)
pls_fit_formula <- plsR(Y ~ ., data = Cornell, nt = 3, pvals.expli = TRUE, verbose = FALSE)

pls_fit_formula$InfCrit
coef(pls_fit_formula)
```

The fitted model stores the extracted components (`tt`), the loadings (`pp`), the coefficients on the original predictors (`Coeffs`), and information-criterion summaries (`InfCrit`).

## Generalized PLS models

```{r glm-fits}
data(aze_compl)
logit_fit <- plsRglm(y ~ ., data = aze_compl, nt = 3, modele = "pls-glm-logistic", verbose = FALSE)

logit_fit$InfCrit
head(predict(logit_fit, type = "response"))

family_fit <- plsRglm(
  Y ~ .,
  data = Cornell,
  nt = 2,
  modele = "pls-glm-family",
  family = gaussian(link = "log"),
  verbose = FALSE
)

family_fit$family$family
family_fit$family$link
```

`plsRglm()` supports predefined model shortcuts together with a custom-family entry point:

```{r supported-modes, eval = FALSE}
plsRglm(Y ~ ., data = Cornell, nt = 3, modele = "pls")
plsRglm(Y ~ ., data = Cornell, nt = 3, modele = "pls-glm-gaussian")
plsRglm(Y ~ ., data = Cornell, nt = 3, modele = "pls-glm-inverse.gaussian")
plsRglm(y ~ ., data = aze_compl, nt = 3, modele = "pls-glm-logistic")
plsRglm(round(x11) ~ ., data = pine, nt = 3, modele = "pls-glm-poisson")
plsRglm(x11 ~ ., data = pine, nt = 3, modele = "pls-glm-Gamma")
plsRglm(Quality ~ ., data = bordeaux, nt = 2, modele = "pls-glm-polr")
plsRglm(
  Y ~ .,
  data = Cornell,
  nt = 3,
  modele = "pls-glm-family",
  family = gaussian(link = "log")
)
```

Ordinal responses are handled through `modele = "pls-glm-polr"`:

```{r polr-fit}
data(bordeaux)
polr_fit <- plsRglm(Quality ~ ., data = bordeaux, nt = 2, modele = "pls-glm-polr", verbose = FALSE)

head(predict(polr_fit, type = "class"))
```

# Cross-Validation and Model Choice

Use `cv.plsR()` for ordinary PLS regression and `cv.plsRglm()` for generalized models. Both provide repeated k-fold cross-validation and integrate with `summary()` and `cvtable()`.

```{r linear-cv}
cv_pls <- cv.plsR(Y ~ ., data = Cornell, nt = 3, K = 4, NK = 2, verbose = FALSE)
cv_pls_summary <- cvtable(summary(cv_pls))

cv_pls_summary
plot(cv_pls_summary)
```

```{r glm-cv}
cv_logit <- cv.plsRglm(
  y ~ .,
  data = aze_compl,
  nt = 3,
  K = 4,
  NK = 2,
  modele = "pls-glm-logistic",
  verbose = FALSE
)
cv_logit_summary <- cvtable(summary(cv_logit, MClassed = TRUE))

cv_logit_summary
plot(cv_logit_summary)
```

For generalized models, `summary(..., MClassed = TRUE)` exposes miss-classification information when it is relevant.

# Prediction and Missing Data

Incomplete predictor matrices are a core package feature, both during fitting and during prediction.

```{r prediction-missing}
data(pine)
data(pine_sup)
data(pineNAX21)

pred_fit <- plsRglm(
  x11 ~ .,
  data = pine,
  nt = 3,
  modele = "pls-glm-family",
  family = gaussian(),
  verbose = FALSE
)

pine_sup_small <- pine_sup[1:3, 1:10]
pine_sup_small[1, 1] <- NA

predict(pred_fit, newdata = pine_sup_small, type = "response", methodNA = "missingdata")
predict(pred_fit, newdata = pine_sup_small, type = "scores", methodNA = "missingdata")

missing_train_fit <- plsR(x11 ~ ., data = pineNAX21, nt = 3, verbose = FALSE)
missing_train_fit$na.miss.X
```

When `newdata` contains incomplete rows, `methodNA = "missingdata"` treats all prediction rows with the missing-data scoring rule, while `methodNA = "adaptative"` switches between complete-row and incomplete-row formulas automatically.

# Bootstrap Utilities

`bootpls()` and `bootplsglm()` wrap the `boot` package for PLS and PLS-GLM models. The default resampling schemes differ:

- `bootpls()` defaults to `(y, X)` resampling with `typeboot = "plsmodel"`.
- `bootplsglm()` defaults to `(y, T)` resampling with `typeboot = "fmodel_np"`.

For a lightweight vignette render, the examples below use a small number of resamples and request non-BCa confidence intervals.

```{r bootstrap}
boot_pls <- bootpls(pls_fit_formula, R = 20, verbose = FALSE)
dim(boot_pls$t)
confints.bootpls(boot_pls, indices = 2:4, typeBCa = FALSE)

boot_logit <- bootplsglm(logit_fit, R = 20, verbose = FALSE)
dim(boot_logit$t)
confints.bootpls(boot_logit, indices = 1:4, typeBCa = FALSE)
```

The plotting helpers `boxplots.bootpls()` and `plots.confints.bootpls()` can be applied directly to these bootstrap objects when a graphical summary is helpful.

# Further Reading

- Use `vignette("plsRglm", package = "plsRglm")` for the historical applications and algorithmic note.
- Use the function help pages for lower-level weighted constructors such as `PLS_lm_wvc()` and `PLS_glm_wvc()`.
- Use `?cv.plsRglm`, `?bootplsglm`, and `?predict.plsRglmmodel` for the full argument reference.

```{r session-information}
sessionInfo()
```
